conmat: generate synthetic contact matrices for a given age-stratified population

### 2024-08-14 paper/paper.Rmd
Fit more word cleanup

There are currently only a few options for a user to access synthetic contact matrices [@socialmixr; @prem]. Existing methods describing how to generate synthetic contact matrices from @prem are not designed for replicability, and are restricted to only selected countries with no sub-national demographic estimates available.

The software exposes model fitting and prediction separately to the user. Users can fit a model based on a contact survey such as POLYMOD, then predict from this model to their own data. This means users can generate synthetic contact matrices for any region, with any contact survey, provided they have the relevant data.

@@

One of the issues with the contact matrices generated by @prem2017 is that some countries are missing. To remedy this we generated synthetic contact matrices for 200 countries, based on a list of country names by the UN, fitted on the POLYMOD contact surveys.
We also ensured that the analysis pipeline is reproducible and transparent by utilising a targets workflow, which allows ease of editing for users.
The resulting synthetic contact matrices, and a replicable / extensible (?) analysis pipeline, can be found in the syncomat analysis pipeline ([GitHub](https://github.com/idem-lab/syncomat), [Zenodo](https://zenodo.org/records/11365943)).

@@

Empirical estimates of social contact are provided by social contact surveys. These provide the frequency and type of social contact across different settings (home, work, school, other).

@@

These social contact surveys can be projected on to a given demographic structure to produce estimated daily contact rates between age groups. These are known as contact matrices or synthetic contact matrices. A widely used approach by @prem2017 [@prem2021] produced contact matrices for 177 countries at "urban" and "rural" levels for each country.

@@

The `conmat` package was developed to fill the specific need of creating contact matrices for arbitrary demographic structures. We developed the method primarily to output contact matrices. We also provided methods to create next generation matrices.

@@

We will generate a contact matrix for Tasmania area within Australia, using a model fitted from the POLYMOD contact survey. We can get the age-stratified population data for Tasmania from the helper function `abs_age_state()`:

@@

Each cell in the resulting contact matrix, indexed *i*, *j*, is the predicted number of people in age group *j* that a single individual in age group *i* will have contact with per day. If you sum across all the *j* age groups for each *i* age group, you get the predicted total number of contacts per day for each individual of age group *i*. [ #TODO expected, predicted, or average? Does it matter? note: currently going with predicted]

@@

The six key features of the relationship between the age groups, represented by the six covariates, are displayed in the figure below.

[ #TODO notes-to-self: the model structure wasn't generated through any particularly robust process, it was just coming up with structures that looked mildly appropriate for our use case. ]

@@

The model structure (covariates?) provided in conmat was designed to represent the key features that are typically present in a contact matrix for different settings (work, school, home, other). Including other sources of information that may better describe these contact patterns, such as ..., may improve model performance.

@@

NOTE: It is also worth noting that there is an updated analysis pipeline that uses conmat, [`syncomat`](https://github.com/idem-lab/syncomat), to generate synthetic contact matrices for 200 countries.

>>>>

There are currently few options for a user to access synthetic contact matrices [@socialmixr; @prem2017]. Existing code to generate synthetic contact matrices from @prem2017 are not designed for replicability, and are restricted to only selected countries with no sub-national demographic estimates available.

The software exposes model fitting and prediction separately to the user. Users can fit a model based on a contact survey such as POLYMOD [@mossong2008], then predict from this model to their own data. This means users can generate synthetic contact matrices for any region, with any contact survey, provided they have the relevant data.

@@

For users who do not wish to run the entire `conmat` pipeline, we have pre-generated synthetic contact matrices for 200 countries, based on a list of country names by the UN, using a model fit to the POLYMOD contact survey. These resulting synthetic contact matrices, and the associated code, can be found in the syncomat analysis pipeline ([GitHub](https://github.com/idem-lab/syncomat), [Zenodo](https://zenodo.org/records/11365943)).

@@

Empirical estimates of social contact are provided by social contact surveys. These provide samples of the frequency and type of social contact across different settings (home, work, school, other).

@@

These social contact surveys can be projected on to a given demographic structure to produce estimated daily contact rates between age groups. These are known as contact matrices or synthetic contact matrices. A widely used approach by @prem2021 [@prem2021] produced contact matrices for 177 countries at "urban" and "rural" levels for each country.

@@

The `conmat` package was developed to fill the specific need of creating contact matrices for arbitrary demographic structures. We developed the method primarily to output synthetic contact matrices. We also provided methods to create next generation matrices.

@@

We will generate a contact matrix for Tasmania, a state in Australia, using a model fitted from the POLYMOD contact survey. We can get the age-stratified population data for Tasmania from the helper function `abs_age_state()`:

@@

Each cell in the resulting contact matrix, indexed $i$, $j$, is the predicted number of people in age group $j$ that a single individual in age group $i$ will have contact with per day. The sum over all of the $j$ age groups for a particular age group $i$ is the predicted total number of contacts per day for each individual of age group $i$.

@@

The six key features of the relationship between the age groups, represented by the six covariates, are displayed in the figure below. These covariates capture typical features of inter-person contact, where individuals primarily interact with people of similar age (the diagonals of the matrix), and with grandparents and/or children (the so-called 'wings' of the matrix).

@@

The model structure (covariates?) provided in conmat was designed to represent the key features that are typically present in a contact matrix for different settings (work, school, home, other). Including other sources of information that may better describe these contact patterns, such as inter-generational mixing, or differences in school ages of a local demographic, may improve model performance.

@@
